---
title: "Homework3"
author: "Siyan Wen"
date: "2023-10-12"
output: github_document
---
```{r}
library(tidyverse)
library(ggridges)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
```
## Problem 1
```{r}
library(p8105.datasets)
data("instacart")
```
**Description:**. 
This dataset contains online grocery data from a online grocery platform called instacart. The size of this dataset is `1384617` rows and `15` columns. There are four variables are `character variables` including evaluation set, product name, aisle and department. The rest of variables are all `integer variables`. In total, it has 15 variables including `order id, product id, customer id, aisle name, department name` and so on. Department name provide general type information of that specific product and then aisle name provide the detail information about which type of product this one belongs to. From the dataset, we can know lots from information. For example, order number 1 contains 4 Bulgarian Yogurt that has its product id to be 49302. This Bulgarian Yogurt belongs to dairy eggs department and also yogurt aisle.  

#### The Total number of Aisles 
```{r}
number_aisle<-
  instacart|>
  count(aisle_id,aisle,sort=TRUE)
```
There are `134` aisle in total within this instacart dataset. `Fresh vagetable` aisle is the most items ordered from.  

#### Plot  
```{r}
number_aisle|>
  filter(n>10000)|>
  ggplot(aes(x=aisle_id,y=n))+
  geom_point()+
  xlim(c(0, 140)) + 
  ylim(c(0, 160000)) + 
  labs(y="Number of Items Ordered", 
       x="Aisle ID", 
       title="Scatterplot of Number of Items Ordered in Each Aisle", 
       )
```
  
From this plot, we can clearly see that there are two specific aisles that most items ordered from.  

#### Table 

```{r}
instacart |> 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) |>
  group_by(aisle) |> 
  count(product_name) |> 
  mutate(rank = min_rank(desc(n))) |> 
  filter(rank < 4) |> 
  arrange(desc(n)) |>
  knitr::kable()
```

#### Table 

```{r}
instacart |>
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) |>
  group_by(product_name, order_dow) |>
  summarize(mean_hour = mean(order_hour_of_day)) |>
  pivot_wider(
    names_from = order_dow, 
    values_from = mean_hour) |>
  knitr::kable(digits = 2)
```

## Problem 2
```{r}
library(p8105.datasets)
data("brfss_smart2010")
```
  
#### Clean the data
```{r}
brfss_smart2010_clean<-
  brfss_smart2010|>
  janitor::clean_names()|>
  filter(topic=="Overall Health")|>
  filter(response==c("Excellent", "Very good", "Good", "Fair", "Poor"))|>
  mutate(response=factor(response, levels = c("Poor", "Fair", "Good", "Very good", "Excellent"), ordered = TRUE))
```

```{r}
brfss_state2002<-brfss_smart2010_clean|>
  filter(year=="2002")|>
  group_by(locationdesc)|>
  summarize(n_obs1=n())|>
  separate(locationdesc,into=c("state", "location"),sep='-')|>
  group_by(state)|>
  summarize(n_obs2=n())|>
  filter(n_obs2>=7)

brfss_state2010<-brfss_smart2010_clean|>
  filter(year=="2010")|>
  group_by(locationdesc)|>
  summarize(n_obs1=n())|>
  separate(locationdesc,into=c("state", "location"),sep='-')|>
  group_by(state)|>
  summarize(n_obs2=n())|>
  filter(n_obs2>=7)
```
In 2002, `no` states were observed at 7 or more locations. In 2010, there are `9` states were oboserved at 7 or more locations.

#### “Spaghetti” plot
```{r}
#Constract dataset
brfss_excellent<-
  brfss_smart2010_clean|>
  filter(response=="Excellent")|>
  select(year,locationabbr,data_value)|>
  group_by(year,locationabbr)|> 
  summarize(mean_data_value=mean(data_value))
#Making plot
ggplot(brfss_excellent,aes(x=year,y=mean_data_value,group=locationabbr,color=locationabbr)) +
  geom_line()+
  xlim(c(2002, 2010)) + 
  ylim(c(5, 35)) + 
  labs(y="Years", 
       x="Mean Data Value", 
       title="“Spaghetti” plot of average data value over time within each state", 
       )
```  

#### Two-panel Plot
```{r}
brfss_two_panal<-
  brfss_smart2010_clean|>
  filter(locationabbr=="NY",year=="2006"|year=="2010")|> 
  select(year,data_value,locationdesc,response)

ggplot(brfss_two_panal, aes(x=response, y=data_value,group=locationdesc,color=locationdesc)) +
geom_line()+
labs(y="Data Value", 
       x="Response", 
       title="Two-panel Plot of Data Value over Responses in 2006 and 2010 among Locations", 
       )+
facet_grid(.~ year) 
```
  
## Problem 3  

#### Clean the dataset
```{r}
nhanes_accel_df= read_csv("data/nhanes_accel.csv")|>
  janitor::clean_names()|>
  drop_na()|>
  mutate(seqn=as.character(seqn))

nhanes_covar_df= read_csv("data/nhanes_covar.csv",skip=4)|>
  janitor::clean_names()|>
  filter(age>=21,)|>
  drop_na()|>
  mutate(seqn=as.character(seqn))|>
  mutate(sex=factor(sex))|>
  mutate(education=factor(education))

nhanes_covar_df$education<-factor(nhanes_covar_df$education, levels=1:3, labels = c("Less than High School", "High School equivalent", "More than High School"))

nhanes_covar_df$sex<-factor(nhanes_covar_df$sex, levels=1:2, labels = c("Male", "Female"))

merged_nhanes <- inner_join(nhanes_covar_df,nhanes_accel_df, by = "seqn")

```  

#### Create a table& Age Distribution
```{r}
sex_education_table<-table(nhanes_covar_df$sex, nhanes_covar_df$education)
sex_education_df <- as.data.frame.matrix(sex_education_table)
print(sex_education_df)
```

```{r}
ggplot(nhanes_covar_df, aes(x = education, y = age, fill = sex)) +
  geom_boxplot() +
  labs(
    title = "Boxplot of Age Distributions for Gender by Education Category",
    x = "Education Category",
    y = "Age"
  ) +
  scale_fill_manual(values = c("Male" = "skyblue", "Female" = "pink")) 
```

Based on the Boxplot, we can see that the age distribution for men and women are similar to each other in the `Less than High School` and `More than High School` Education Category. However, in the `High School equivalent` category, female are relatively older than male based on this dataset. 
  
#### Total Activity Plot
```{r}
merged_nhanes$total_activity<-rowSums(merged_nhanes[,6:1445])
ggplot(merged_nhanes, aes(x = age, y = total_activity,group=sex, color=sex)) +
  geom_point()+
  geom_smooth() +
  labs(
    title = "Total Activities vs Age",
    x = "Age",
    y = "Total Activity"
  ) +
  facet_grid(.~ education) 
```   

Based on this plot, the total activity decrease when people geting older and older across all three education category for both men and women. For people who received `less than high school` education, they have two high total activity peroids in their lifetime. One of them locate at around 20 years old and the other peak at around 58 years old. Women have higher total activity than men earlier in their life and men have their higher activity than women later in life.For people who received `High School equivalent` education, they have only one total activity peak during their life. This one big peak located at around 40 years old for both men and women. For `More than High School` category, their total activity over their lifetime is relatively stable. Generally speaking, women have higher total acvity than men for `High School equivalent` and `More than High School` category.   

#### 24-hour Activity Plot
```{r}

minutes <- seq(from = 1, to = 1440, by = 1)

new_nhanes_accel_df<-
  pivot_longer(nhanes_accel_df,min1:min1440,
               names_prefix = "min",
               names_to = "minutes",
               values_to = "activity")
nhanes <- inner_join(nhanes_covar_df,new_nhanes_accel_df,by = "seqn")

ggplot(nhanes, aes(x = minutes , y = activity, group=sex, color=sex)) +
  geom_line(aes(color = sex),alpha=0.2)+
  geom_smooth(aes(color = sex), se = FALSE, linetype = "longdash") +
  labs(
    title = "Activities vs Minutes",
    x = "Minutes",
    y = "Activities"
  )+
  scale_color_manual(values = c("Male" = "dodgerblue", "Female" = "deeppink"))+
  facet_grid(.~ education)
```

From the smooth line, we can see that both women and men shows similar patterns among three different education categories. However, based on the line, we can see that male from higher eductaion level tends to have more activities in the morning, and women from higher education level tends to have more activities in the afternoon. 






